---
title: "W4 Math - Linear Algebra"
author: "Kyle Gilde"
date: "December 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## W4 Math - Linear Algebra
### Homework

1. Using matrix operations, describe the solutions for the following family of equations:
x + 2y - 3z = 5
2x + y - 3z = 13
-x + y + 2z= -8
a.  Find the inverse of the above 3x3 (non-augmented) matrix.
```{r}
D <-matrix(c(1,2,-3,5,2,1,-3,13,-1,1,2,-8),ncol=4,byrow=T)
A <- D[,1:3] 
B <- D[,4]
solve(A)
```
b.  Solve for the solution using R.
```{r}
solve(A, B)
```
c.  Modify the  3x3 matrix such that there exists only one non-zero variable in the solution set.
```{r}
E <- A
E[,2] <- B
E
solve(E, B)
```
2.  Consider the matrix, q=matrix(c(3,1,4,4,3,3,2,3,2),nrow=3). Let b=c(1,4,5).  Use Cramer's rule and R to determine the solution, x, to qx=b, if one exists.  Show all determinants.
```{r}
q <- matrix(c(3,1,4,4,3,3,2,3,2), nrow=3)
b <- c(1,4,5)

D <- det(q) #coefficient determinant
Dx <- det(cbind(b, q[,2:3]))
Dy <- det(cbind(q[,1], b, q[,3]))
Dz <- det(cbind(q[,1:2], b))

D;Dx;Dy;Dz
solve(q, b)
```


### Introductory L. A. w/ R

```{r}

a <- c(2, 1, 3)
b <- c(3, 4, 5)
t(a)

#Inner Product
sum(a*b)

# 1.2.8 Orthogonal (perpendicular) vectors
# Two vectors v1 and v2 are orthogonal if their inner product is zero, written
# v1 ⊥ v2 ⇔ v1 · v2 = 0
# Note that any vector is orthogonal to the 0–vector. Orthogonal vectors:
v1 <- c(1,1)
v2 <- c(-1,1)
sum(v1*v2)

#Matrices
A <-  matrix(c(1,3,2,2,8,9),ncol=3)
A2 <- matrix(c(1,3,2,2,8,9),ncol=3,byrow=T)
A
A2

#Vector & Matrix Multiplication
A <- matrix(c(1,2,3,3,2,1),ncol=3, byrow = T) #2 by 3
a <- c(3, 2, 1) # 3 by 1 
A
as.matrix(a) # a vector is treated like a 1-column matrix
A%*%a # 2 by 1 
A*a #this seems to cycle through the vector elements until all matrix elements have been multiplied

#Special Matrices
matrix(0,nrow=2,ncol=3)
matrix(1,nrow=2,ncol=3)
diag(c(1,2,3))
diag(1,3)

diag(diag(c(1,2,3)))
diag(A)

# Some facts about inverse matrices are:
# • Only square matrices can have an inverse, but not all square matrices have an inverse.
# • When the inverse exists, it is unique.
# • Finding the inverse of a large matrix A is numerically complicated (but computers do
# it for us).
D <- matrix(c(1,3,2,4),ncol=2,byrow=T)
solve(D)%*%D

#Solving systems of linear equations: If Mx = z where M is a matrix and x and z are vectors the solution is x = M−1z:
A <- matrix(c(1,2,3,4),ncol=2)
b <- c(7,10)
x <- solve(A, b)
A
b
x

```



### Slides
Properties of Matrix Addition
1. Commutative Property
A + B = B + A (Order does not matter)
2. Associative Property
(A + B) + C = A + (B + C) (Grouping does
not matter)
3. Additive Identity
A + 0 = 0 + A = A (When you add to 0 then the
matrix keeps its identity)
4. Additive Inverse
A + (-A) = -A + A = 0 (If you add the
opposite, then you get 0.)

```{r Slides Code & Notes}
rm(list = ls())
A=matrix(c("a11","a21","a12","a22"),2,2)
B=matrix(c(1,5,-2,3,9,7,-4,0,6),3,3)
C=matrix(c(-4,0,5,3,0,-2),2,3)
D=matrix(c(1,3,-2,5,0,4),2,3)
E=matrix(c(4,-2,3,2,-1,-4),2,3)
D * E
D+E
F1=matrix(c(5,9,7,-6,10,-2),3,2)
G=matrix(c(5,-3,6,6,11,-1),3,2)
F1-G

#Properties of Matrix Addition
F+G
G+F
(C+D)+E
C+(D+E)
B+0 #R knows that this is a matrix of zeros
0+B
B+-B
-B+B

#Scalar Multiplication
H=matrix(c(-3,0,6,-9),2,2)
l=2
l*H
#Example:
J=matrix(c(5,6,0,-2),2,2)
K=matrix(c(0,3,-1,5),2,2)
m=2
n=-3
m*J+n*K

#Multiplying Matrices
#Rows A * Columns of B, inner dimensions must match!!
#2 x 2 x 2 x 1
L=matrix(c(1,0,2,4),2,2)
M=matrix(c(5,-2),2,1)
L
M
L%*%M
#what is M%*%L…???
M%*%L
#What about t(M)%*%L  Error in M %*% L : non-conformable arguments
M
t(M)
t(M)%*%L
# Multiplying Matrices Ex 2
N=matrix(c(2,0,-8,4),2,2)
O=matrix(c(1,-4,3,5,0,-2),2,3)
N%*%O
#Identity Matrix
#Finding the multiplicative Inverse of a 2x2 Matrix

A=matrix(c(5,-3,0,2),2,2)
B=matrix(c(2,3,0,5),2,2)
solve(A)
#or
1/det(A)*B

#Birdie table problem
B <- matrix(c(1,-1,-2,1,-.5,1.5),ncol=3,byrow=T)
solve(B[,1:2],B[,3])

fx <- function(x) .61 * x + 3.07
fx(1:5)
```

Steps to finding the multiplicative inverse
1. Form the augmented matrix [A│In]
2. Perform row operations on it to obtain a matrix of the form [In│B].
3. Matrix B is the inverse matrix.
4. Verify the results by multiplying AB to see if you get In
Matrix Row Operations
1. Two rows may be interchanged
2. The elements in any row may be multiplied by a nonzero number
3. The elements in any row may be multiplied by a nonzero number, and
these products may be added to the corresponding elements in any other row.
 